### 2020-06-30 Report/aws_325ghz_study.tex
updated with comments from Christophe

\section{Introduction}
%
\subsection{Background}
The Arctic Weather Satellite (AWS) is a small platform carrying a single
radiometer package. This instrument is an across-track scanning microwave
radiometer. The exact channel configuration of AWS is not yet set. The main
issue for this study is that AWS will have channels at 183\,GHz, but also one
or several channels a higher frequencies. A number of channels around 325\,GHz
was considered as the main option before AWS became an ESA/EUMETSAT mission,
but it has now also been suggested to have a single channel at 229\,GHz. This
suggestion is based on MWS (MicroWave Sounder), one of the instruments onboard
the next generation of Metop satellites. The 229\,GHz was added to MWS as an
help for performing cloud screening of 183\,GHz data
\citep{rekha2012potential}. For comparison, it can be mentioned that ICI (Ice
Cloud Imager), another new Metop sensor, will include 325\,GHz channels.

@@

\subsubsection{Bulk profile database}
%
A database of atmospheric cases was created based on data from CloudSat and
ERA-Interim. The core idea of the approach is to use CloudSat reflectivities to
obtain as realistic vertical profiles of ice hydrometeors and rain as
possible. No external CloudSat retrievals are involved. Instead the CloudSat
reflectivities are mapped to properties for the passive simulations by assuming
a particle size distribution and a particle habit, together denoted as
the particle model. The process involves an inversion of the radar data, but as,
the same particle model is used for that inversion and to map ice water content
(IWC) and rain water content (RWC) to single scattering data when performing
the passive radiative transfer calculations, the impact of clouds on the
simulated AWS data is as realistic as possible. The thin clouds not detected by
CloudSat should not be of relevance for passive measurements frequencies
350\,GHz. Background data (temperature, humidity, 10\,m wind speed, \dots) were
taken from ERA-Interim, for the time and location of the CloudSat data. Liquid
water content (LWC) was taken from ERA-Interim, as CloudSat has no sensitivity
to the hydrometeor category.

@@

For RWC the particle size distribution of \citet{abel2012improved} is applied.
The PSD of IWC follows the basic formulation applied in DARDAR, using latest
parameter values (i.e.\ $\alpha$ and $\beta$) as given by
\citet{amt-12-2819-2019}. This PSD can be considered as a ``two moment''
scheme, but is here applied in a one moment manner by setting $N_0^*$ (as a
function of temperature) following Table~5 of \citet{delanoe2014normalized},
and letting the radar reflectivity set the remaining moment.

@@

\subsection{Validation of simulations}
\label{sec:validation}
%
\begin{figure}[p]
	\centering
	\includegraphics[height=60mm]{ATMS_C16_distribution}\hspace{5mm}% 
	\includegraphics[height=60mm]{ATMS_C17_distribution}
	\includegraphics[height=60mm]{ATMS_C18_distribution}\hspace{5mm}% 
	\includegraphics[height=60mm]{ATMS_C19_distribution}
	\includegraphics[height=60mm]{ATMS_C20_distribution}\hspace{5mm}% 
	\includegraphics[height=60mm]{ATMS_C21_distribution}
	\includegraphics[height=60mm]{ATMS_C22_distribution}
	\caption{Probability distribution functions (PDFs) of simulated and
      observed ATMS channels brightness temperatures for channels 16 - 22 and
      nadir viewing angle. The data covers latitude range $60^{\degree}$S to
      $60^{\degree}$N. Both land and ocean measurements are considered and only
      daytime ATMS data are included. See further the text.}
	\label{fig:pdf:c16-22}
\end{figure}

@@

The overall agreement is good for all channels. For 183 GHz channels (Channels
18-22), the simulations and observations have a good agreement for the part of
the PDFs matching clear-sky conditions (the main peak of the distribution),
albeit the simulations have a more sharp peak. This could either be caused by a
different geographical sampling of ATMS and CloudSat or to shortcomings in the
ERA-Interim input data. The fraction of cloudy observations is lower in ATMS in
comparison to simulations, for brightness temperatures below about 240\.K. This
is due to the fact that the antenna pattern is not considered in the
simulations.

@@

The altitude variation of transmissivity exhibits same basic shape between all
the 183 and 325\,GHz channels, but the altitude of optical thickness of 1
(transmissivity of $e^{-1}\approx0.37$) varies. The curves of the edge 183\,GHz
channels (AWS-32 and 36) brackets the ones corresponding to the ICI 325\,GHZ
channels. The later channels show a higher deviation from 1 at 150\,hPa, due to
some (weak) ozone attenuation in the stratosphere. The 229\,GHz has the highest
transmissivities.

@@

\subsection{A three channel option}
%
The ICI channels give a rough RH weighting functions matching with AWS-32,
AWS-33 and AWS-34 (Fig.~\ref{fig:wf:mls:025}), and to find a tentative three
channel option an even better matching was used as constrain to set the
specifications. The resulting values are found in Table~\ref{tab:chs:prel3} and
example weighting functions are displayed in Fig.~\ref {fig:3ch:prel}. This
option has a IF range of 1.0 to 8.0\,GHz, and leaves a 100\,MHz space between
AWS-42 and AWS-43, and 200\,MHz space between AWS-41 and AWS-42. That is, the
suggestion uses the complete IF bandwidth as far as possible, that is
technically possible.

\subsection{A four channel option}
%
The tentative four channel option in Table~\ref{tab:chs:prel4} gives a broader
altitude coverage, towards higher altitudes, and offers weighting functions
roughly equally spaced in altitude (Fig.~\ref {fig:4ch:prel}). This option has
a IF range of 0.7 to 8.0\,GHz, and leaves 200\,MHz space between all the
325\,GHz channels. Also this suggestion uses the complete IF bandwidth.

@@

\subsection{A second three channel option}
%
Relatively late in the study Omnisys informed us about that an IF of 1.7\,GHz
should be avoided, as this frequency likely will be used for downlink of data.
The four channel option needed just a marginal change to adopt to this
constrain, while 1.7\,GHz is more or less at the centre of the three channel
option. It was then decided to formulate second three channel option avoiding 
1.7\,GHz, and the suggestion is found in Table~\ref{tab:chs:3b}.

@@

\section{Cloud correction by a simple scheme}
\label{sec:simple}
%
In this section, the (first) 325\,GHz three channel option is evaluated for detection
and correction of cloud interference in 183\,GHz channels. Combinations of
brightness temperature (TB) differences between 183\,GHz and 325\,GHz are used
as a measure of the cloud impact. A regression based approach is used to
estimate the adjustment needed to predict the clear-sky values.

@@

We will simply treat cases that matches a correction below a certain value as
``clear-sky''. If the correction limit is denoted as $dtb_{cs}$, the threshold
value, $x_{cs}$, is obtained by inverting the polynomial fit:
\begin{equation}
x_{cs} = f_{cs}^{-1}(dtb_{cs}).
\label{eq:dtb}
\end{equation}
That is, if $x_m>x_{cs}$ a measurement is classified as clear-sky. If
$x_m\leq x_{cs}$, the measurement is rejected if a pure filtering is made, or
it is adjusted if the correction scheme is applied. We will largely set
$dtb_{cs}$ based on the noise level of the 183\,GHz channel involved, to be
1$\sigma$ or 2$\sigma$ where $\sigma$ refers to the magnitude of measurement
noise (Eq.~\ref{eq:noise}, often called NE$\Delta$T). We will use $1\sigma$ for
the development of the scheme, however, the sensitivity of the scheme to
a 1$\sigma$ or 2$\sigma$ threshold is also provided later in the text.

@@

The distributions of filtered data are non-Gaussian. The right side of the
distribution follows measurement noise while the left side is skewed. The
skewness in the distribution is due to cloud impacts which pass through the
clear-sky threshold. The skewness in ``Sreerekha et al'' is higher than our approach because their method only aims at removing cloud impacts above 4\,K. Their approach also leads to a high rejection fraction (46.37\%). The filtering
method introduced here leads to a considerable lower rejection of data
(17.54\%).

@@

\subsubsection{Training}
%  
\label{sec:qrnn:training}
\begin{figure}[!tb]
	\centering
	\includegraphics[height=85mm]{quantile_loss.png}
	\caption{The mean quantile loss with different configurations of layer width and hidden layers ($n_h$). Here QRNN is trained to predict clear-sky values for AWS-34 using four channels option of 325\,GHz}
	\label{fig:quantile_loss}
\end{figure}
In this section, QRNNs are used to correct the cloud impact in each of 183\,GHz
channels using TBs measurements from all available channels of 325\,GHz. A
separate QRNN is used for each of the 183\,GHz channels. For example, to
predict clear-sky TBs from AWS-34, the training data is measurements from all
channels of 325\,GHz and measurements from the 183\,GHz to be predicted. We
assess the performance of the scheme using both three channel and four channel
options of 325\,GHz. Two suggestions for three channel option are described in
Sec.~\ref{sec:channel:configurations}, and both have been used to train QRNN
separately. Thus, for each channel of 183\,GHz, we have a set of three
predictions: one from four channel option, and two from the old and new
configurations of the three channel option. Further in the text, the old and
new three channel configurations are labelled as ``three-a'' and ``three-b''.

@@

At this point we got a feedback from Omnisys that also channel option three-b is
technically challenging. In short, it is significantly more difficult to
implement a 3\,GHz wide channel at an IF of 3.3\,GHz (AWS-42) than compared to
at 6.5\,GHz (AWS-41). To decrease the ratio between channel widths and centre
IF, a third three channel option was formulated (Table~\ref{tab:chs:3c}) that
is below denoted as three-c. This option uses the same IF range as the
previously suggested ones.

@@

\subsection{Impact of extending the IF range}
%
As a basic investigation of the impact of changing the width of the IF range, a
version of the four channel option was considered, denoted as four-b. It was
modified to cover 0.7 to 9.0 GHz (instead of 0.7 to 8.0 GHz). To focus on the
extension, only the edge channels were changed. That is, the revised AWS-41 and AWS-44
have bandwidth of 3.8 and 0.9\,GHz, respectively. The DoFs obtained for four-b
are found in Table~\ref{tab:DoF:4b}.

>>>>

\section{Introduction}
%
\subsection{Background}
The Arctic Weather Satellite (AWS) is a small platform carrying a single
radiometer package. This instrument is an across-track scanning microwave
radiometer. The exact channel configuration of AWS is not yet set. The main
issue for this study is that AWS will have channels at 183\,GHz, but also one
or several channels a higher frequencies. A number of channels around 325\,GHz
was considered as the main option before AWS became an ESA prototype mission,
but it has now also been suggested to have a single channel at 229\,GHz. This
suggestion is based on MWS (MicroWave Sounder), one of the instruments onboard
the next generation of Metop satellites. The 229\,GHz was added to MWS as an
help for performing cloud screening of 183\,GHz data
\citep{rekha2012potential}. For comparison, it can be mentioned that ICI (Ice
Cloud Imager), another new EPS-SG (EUMETSAT Polar System - Second Generation) sensor, will include 325\,GHz channels.

@@

\subsubsection{Bulk profile database}
%
A database of atmospheric cases was created based on data from CloudSat \citep{stephens2002bams} and
ERA-Interim \citep{dee2011qjrms}. The core idea of the approach is to use CloudSat reflectivities to
obtain as realistic vertical profiles of ice hydrometeors and rain as
possible. No external CloudSat retrievals are involved. Instead the CloudSat
reflectivities are mapped to properties for the passive simulations by assuming
a particle size distribution and a particle habit, together denoted as
the particle model. The process involves an inversion of the radar data, but as,
the same particle model is used for that inversion and to map ice water content
(IWC) and rain water content (RWC) to single scattering data when performing
the passive radiative transfer calculations, the impact of clouds on the
simulated AWS data is as realistic as possible. The thin clouds not detected by
CloudSat should not be of relevance for passive measurements frequencies
350\,GHz. Background data (temperature, humidity, 10\,m wind speed, \dots) were
taken from ERA-Interim, for the time and location of the CloudSat data. Liquid
water content (LWC) was taken from ERA-Interim, as CloudSat has no sensitivity
to the hydrometeor category.

@@

For RWC the particle size distribution of \citet{abel2012improved} is applied.
The PSD of IWC follows the basic formulation applied in DARDAR (\verb http://www.icare.univ-lille1.fr/projects/dardar), using latest
parameter values (i.e.\ $\alpha$ and $\beta$) as given by
\citet{amt-12-2819-2019}. This PSD can be considered as a ``two moment''
scheme, but is here applied in a one moment manner by setting $N_0^*$ (as a
function of temperature) following Table~5 of \citet{delanoe2014normalized},
and letting the radar reflectivity set the remaining moment.

@@

\subsection{Validation of simulations}
\label{sec:validation}
%

\begin{figure}[p]
	\centering
	\includegraphics[height=57mm]{ATMS_C16_distribution}\hspace{5mm}% 
	\includegraphics[height=57mm]{ATMS_C17_distribution}
	\includegraphics[height=57mm]{ATMS_C18_distribution}\hspace{5mm}% 
	\includegraphics[height=57mm]{ATMS_C19_distribution}
	\includegraphics[height=57mm]{ATMS_C20_distribution}\hspace{5mm}% 
	\includegraphics[height=57mm]{ATMS_C21_distribution}
	\includegraphics[height=57mm]{ATMS_C22_distribution}
	\caption{Probability distribution functions (PDFs) of simulated and
      observed ATMS channels brightness temperatures for channels 16 - 22 and
      nadir viewing angle. The data covers latitude range $60^{\degree}$S to
      $60^{\degree}$N. Both land and ocean measurements are considered and only
      daytime ATMS data are included. See further the text.}
	\label{fig:pdf:c16-22}

\end{figure}

@@

The overall agreement is good for all channels. For 183 GHz channels (Channels
18-22), the simulations and observations have a good agreement for the part of
the PDFs matching clear-sky conditions (the main peak of the distribution),
albeit the simulations have a more sharp peak. This could either be caused by a
different geographical sampling of ATMS and CloudSat or to shortcomings in the
ERA-Interim input data. The fraction of cloudy observations is lower in ATMS in
comparison to simulations, for brightness temperatures below about 240\,K. This
is due to the fact that the antenna pattern is not considered in the
simulations.

@@

The altitude variation of transmissivity exhibits same basic shape between all
the 183 and 325\,GHz channels, but the altitude of optical thickness of 1
(transmissivity of $e^{-1}\approx0.37$) varies. The curves of the edge 183\,GHz
channels (AWS-32 and 36) bracket the ones corresponding to the ICI 325\,GHZ
channels. The later channels show a higher deviation from 1 at 150\,hPa, due to
some (weak) ozone attenuation in the stratosphere. The 229\,GHz has the highest
transmissivities.

@@

\subsection{A three channel option}
%
The ICI channels give a rough RH weighting functions matching with AWS-32,
AWS-33 and AWS-34 (Fig.~\ref{fig:wf:mls:025}), and to find a tentative three
channel option an even better matching was used as constrain to set the
specifications. The resulting values are found in Table~\ref{tab:chs:prel3} and
example weighting functions are displayed in Fig.~\ref {fig:3ch:prel}. This
option has a IF range of 1.0 to 8.0\,GHz, and leaves a 100\,MHz space between
AWS-42 and AWS-43, and 200\,MHz space between AWS-41 and AWS-42. That is, the
suggestion uses the complete IF bandwidth as far as possible, that is
technically possible. Further in the text, this channel option is denoted as ``three-a''.

\subsection{A four channel option}
%
The tentative four channel option in Table~\ref{tab:chs:prel4} gives a broader
altitude coverage, towards higher altitudes, and offers weighting functions
roughly equally spaced in altitude (Fig.~\ref {fig:4ch:prel}). This option has
a IF range of 0.7 to 8.0\,GHz, and leaves 200\,MHz space between all the
325\,GHz channels. Also this suggestion uses the complete IF bandwidth. We will denote this channel option as ``four''.

@@

\subsection{A second three channel option}
%
Relatively late in the study Omnisys informed us about that an IF of 1.7\,GHz
should be avoided, as this frequency likely will be used for downlink of data.
The four channel option needed just a marginal change to adopt to this
constrain, while 1.7\,GHz is more or less at the centre of the three channel
option. It was then decided to formulate a second three channel option, denoted as ``three-b'', avoiding 1.7\,GHz, and the suggestion is found in Table~\ref{tab:chs:3b}.

@@

\section{Cloud correction by a simple scheme}
\label{sec:simple}
%
In this section, the (first) 325\,GHz three channel option (three-a) is evaluated for detection
and correction of cloud interference in 183\,GHz channels. Combinations of
brightness temperature (TB) differences between 183\,GHz and 325\,GHz are used
as a measure of the cloud impact. A regression based approach is used to
estimate the adjustment needed to predict the clear-sky values.

@@

We will simply treat cases that matches a correction below a certain value as
``clear-sky''. If the correction limit is denoted as $dtb_{cs}$, the threshold
value, $x_{cs}$, is obtained by inverting the polynomial fit:
\begin{equation}
x_{cs} = f_{cs}^{-1}(dtb_{cs}).
\label{eq:dtb}
\end{equation}
That is, if $x_m>x_{cs}$ a measurement is classified as clear-sky. If
$x_m\leq x_{cs}$, the measurement is rejected if a pure filtering is made, or
it is adjusted if the correction scheme is applied. We will largely set
$dtb_{cs}$ based on the noise level of the 183\,GHz channel involved, to be
1$\sigma$ or 2$\sigma$ where $\sigma$ refers to the magnitude of measurement
noise (Eq.~\ref{eq:noise}, often called NE$\Delta$T (Noise Equivalent Delta Temperature)). We will use $1\sigma$ for
the development of the scheme, however, the sensitivity of the scheme to
a 1$\sigma$ or 2$\sigma$ threshold is also provided later in the text.

@@

The distributions of filtered data are non-Gaussian. The right side of the
distribution follows measurement noise while the left side is skewed. The
skewness in the distribution is due to cloud impacts which pass through the
clear-sky threshold. The skewness in ``Sreerekha et al'' is higher than our approach because their method only aims at removing cloud impacts above 4\,K. Their approach also leads to a high rejection fraction (46.37\%). The filtering 
method introduced here, rejects only 17.54\% of the data.

@@

\subsubsection{Training}
%  
\label{sec:qrnn:training}
\begin{figure}[!tb]
	\centering
	\includegraphics[height=85mm]{quantile_loss.png}
	\caption{The mean quantile loss with different configurations of layer width and hidden layers ($n_h$). Here QRNN is trained to predict clear-sky values for AWS-34 using four channels option of 325\,GHz}
	\label{fig:quantile_loss}
\end{figure}
In this section, QRNNs are used to correct the cloud impact in each of 183\,GHz
channels using TBs measurements from all available channels of 325\,GHz. A
separate QRNN is used for each of the 183\,GHz channels. For example, to
predict clear-sky TBs from AWS-34, the training data is measurements from all
channels of 325\,GHz and measurements from the 183\,GHz to be predicted. We
assess the performance of the scheme using both three channel and four channel
options of 325\,GHz. Two suggestions for three channel option are described in
Sec.~\ref{sec:channel:configurations}, and both have been used to train QRNN
separately. Thus, for each channel of 183\,GHz, we have a set of three
predictions: one from four channel option, and two from the options ``three-a'' and ``three-b''.

@@

At this point we got a feedback from Omnisys that also channel option three-b is
technically challenging. In short, it is significantly more difficult to
implement a 3\,GHz wide channel at an IF of 3.3\,GHz (AWS-42) than compared to
at 6.5\,GHz (AWS-41). To decrease the ratio between channel widths and centre
IF, a third three channel option was formulated (Table~\ref{tab:chs:3c}) that
is below denoted as ``three-c''. This option uses the same IF range as the
previously suggested ones.

@@

\subsection{Impact of extending the IF range}
%
As a basic investigation of the impact of changing the width of the IF range, a
version of the four channel option was considered, denoted as ``four-b''. It was
modified to cover 0.7 to 9.0 GHz (instead of 0.7 to 8.0 GHz). To focus on the
extension, only the edge channels were changed. That is, the revised AWS-41 and AWS-44
have bandwidth of 3.8 and 0.9\,GHz, respectively. The DoFs obtained for four-b
are found in Table~\ref{tab:DoF:4b}.
